An examination of the high-order dynamic interactions underlying multi-dimensional timeseries data

### 2019-03-27 paper/main.tex

\subsubsection*{Recovery of ground truth parameters from synthetic
  data}
We applied timecorr with a given kernel Fig.~\ref{fig:weights}) to
synthetic data, then correlate each recovered
correlation matrix with the ground truth.  Explore how recovery varies
with the kernel, kernel parameters, and specific structure of the data
(e.g. slow changes as in the ramping dataset, versus rapid changes as
in the block dataset).

@@

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
For our decoding analysis, we used HTFA-derived node activities
~\cite{MannEtal18} from fMRI data collected as participants listened
to an audio recording of a story (intact condition; 36 participants),
listened to time scrambled recordings of the same story (17
participants in the paragraph-scrambled condition listened to the
paragraphs in a randomized order and 36 in the word-scrambled
condition listened to the words in a randomized order), or lay resting
with their eyes open in the scanner (rest condition; 36
participants). We sought to demonstrate how higher-order correlations
may be used to examine dynamic interactions of brain patterns in
(real) multi-subject fMRI datasets. This story listening dataset was
collected as part of a separate study, where the full imaging
parameters, image preprocessing methods, and experimental details may
be found~\citep{SimoEtal16}. The dataset is available at
\href{url}{http://arks. princeton.edu/ark:/88435/dsp015d86p269k}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Decoding by order.
      a. Relative decoding accuracy by order.} Relative to chance,
    decoding accuracy for each condition (intact, paragraph, word,
    and rest) as a function of increasingly more orders.  The red dots indicates
    maximum decoding accuracy for each condition. \textbf{b. Normalized decoding accuracy
      by order.}  We normalized the decoding accuracy by order to better
    visualize the order with the maximum decoding accuracy for each
    condition. \textbf{c. Optimized weights.}
  For the order with maximum decoding accuracy by condition, we show
  barplots of the optimized weights $\phi$  for each contributing
  order.}
  \label{fig:decoding_level}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/corrmats.pdf}
  \caption{\textbf{Inter-subject correlations by order.
      a. Inter-subject correlation heatmaps by order.}  Heatmap
    denoting how correlated each subject is with every other subject
    for each order, and grouped by condition (intact (I), paragraph
    (P), word (W),
    and rest (R)). \textbf{b. Distribution of correlations for each
      condition by order}
   Inter-subject correlations by conditon for each order.}
  \label{fig:corrmats}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
\begin{itemize}
\item multiple timescale representations (a la Hasson group) implies
first-order network interactions.  higher-order interactions imply
generalizations between interacting representations (e.g. mirrored
schema, a la Norman/Baldassano/Hasson).  possibly cite NTB 2013
science review, using as evidence that this is where the field is
going (voxels $\rightarrow$ patterns (L0) $\rightarrow$ interactions (L1) $\rightarrow$ higher
order patterns (L2$+$).
\item related approaches: sliding window, phase-based correlations,
within-ROI spatial correlations at each timepoint, granger
causality, other explit models (e.g. virtual brain).
\item other applications: molecular interactions (protein folding?),
diagnosis (e.g. psychiatric disorders as network flow problems--
gratton work?), social network dynamics (e.g. financial markets,
social media interactions)
\end{itemize}

>>>>

\subsubsection*{Recovery of ground truth parameters from synthetic
  data}

To explore recovery of a constant correlation, we created a random
correlation matrix and and added a small amount of Gaussian noise to

We applied timecorr with a given kernel Fig.~\ref{fig:weights}) to
synthetic data, then correlate each recovered
correlation matrix with the ground truth.  Explore how recovery varies
with the kernel, kernel parameters, and specific structure of the data
(e.g. slow changes as in the ramping dataset, versus rapid changes as
in the block dataset).

@@

\subsubsection*{Neuroimaging dataset~\citep{SimoEtal16}}
For our decoding analysis, we used HTFA-derived node activities
~\cite{MannEtal18} from fMRI data collected as participants listened
to an audio recording of a story (intact condition; 36 participants),
listened to time scrambled recordings of the same story (17
participants in the paragraph-scrambled condition listened to the
paragraphs in a randomized order and 36 in the word-scrambled
condition listened to the words in a randomized order), or lay resting
with their eyes open in the scanner (rest condition; 36
participants). We sought to demonstrate how higher-order correlations
may be used to examine dynamic interactions of brain patterns in
(real) multi-subject fMRI datasets. This story listening dataset was
collected as part of a separate study, where the full imaging
parameters, image preprocessing methods, and experimental details may
be found~\citep{SimoEtal16}. The dataset is available at
\href{url}{http://arks. princeton.edu/ark:/88435/dsp015d86p269k}.
Bars of each color display cross-validated decoding performance for decoders trained using different sets of neural features: whole-brain patterns of voxel activities
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Decoding by order.
      a. Relative decoding accuracy by order.} Ribbons of each
    color display cross-validated decoding performance for each condition (intact, paragraph, word,
    and rest. Decoders were trained using increasingly more
    higher-order information and this ribbons are displayed relative
    to chance at 0. The red dots indicates
    maximum decoding accuracy for each condition. \textbf{b. Normalized decoding accuracy
      by order.}  We normalized the decoding accuracy by order to better
    visualize the order with the maximum decoding accuracy for each
    condition. \textbf{c. Optimized weights.} Bar heights indicate the
    optimized mixing paramete $\phi$ of each contributing order up to
    and including the order with the maximum
    decoding accuracy for each contributing order
  For the order with maximum decoding accuracy by condition, we show
  barplots of the optimized weights $\phi$  for each contributing
  order.}
  \label{fig:decoding_level}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/corrmats.pdf}
  \caption{\textbf{Inter-subject correlations by order.
      a. Inter-subject correlation heatmaps by order.}  Heatmaps for
    each order of how correlated each subject is with every other
    subject grouped by condition (intact (I), paragraph
    (P), word (W),
    and rest (R)). \textbf{b. Mean correlation for each
      condition by order.}
   Bar heights indicate inter-subject correlations by conditon for each order.}
  \label{fig:corrmats}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
\begin{itemize}

\item Decoding accuracy best for level 2 data for all but rest
  condition.  Could be that the brain is 2nd order or that fMRI can
  only reliable give 2nd order.  
\item multiple timescale representations (a la Hasson group) implies
first-order network interactions.  higher-order interactions imply
generalizations between interacting representations (e.g. mirrored
schema, a la Norman/Baldassano/Hasson).  possibly cite NTB 2013
science review, using as evidence that this is where the field is
going (voxels $\rightarrow$ patterns (L0) $\rightarrow$ interactions (L1) $\rightarrow$ higher
order patterns (L2$+$).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/direction_of_field.pdf}
  \caption{\textbf{Direction of the field.} }
  \label{fig:direction_of_field}
\end{figure}

\item related approaches: sliding window, phase-based correlations,
within-ROI spatial correlations at each timepoint, granger
causality, other explit models (e.g. virtual brain).
\item other applications: molecular interactions (protein folding?),
diagnosis (e.g. psychiatric disorders as network flow problems--
gratton work?), social network dynamics (e.g. financial markets,
social media interactions)
\end{itemize}
